Negative and Fractional Types

### 2012-06-27 rational_ext.tex
_
started on negatives

% \dgm{if-c1-c2}

@@

%%%%%%%%%
\subsection{Categorical Constructions}
\label{sec:frac-constructions}

All the constructions below are standard: they are collected from Selinger's
survey paper on monoidal
categories~\cite{springerlink:10.1007/978-3-642-12821-94} and presented in
the context of our language.

@@

We illustrate the expressive power of first-class constraints represented by
fractional types. To understand the intuition, recall the definition of
{{trace}} for the multiplicative fragment of {{PiEE}}. 
Given {{f : a*c <-> b*c }}, we have {{trace* f : a <-> b}}:

A large class of constraint satisfaction problems can be expressed
using {{trace*}}. We illustrate the main ideas with the implementation
of a SAT-solver.  We proceed in small steps, reviewing some of the
necessary constructions presented in our earlier
paper~\cite{infeffects}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Finite Dimensional Hilbert Spaces}

%subcode{opsem} include main
%! columnStyle = rclr
%! bck = \triangleleft
% {[iso, v, C, s]}_{bck} &|-->& <iso, v', C, s'>_{bck}
% & & where iso^{dagger} s v |--> (v', s')
% <c1, v, Fst C c2, s>_{bck} &|-->& <c1(;)c2, v, C, s>_{bck}
% <c2, v, Snd c1 C, s>_{bck} &|-->& {[c1, v, Fst C c2, s]}_{bck}
% {[ c1(;)c2, v, C, s ]}_{bck} &|-->& {[c2, v, Snd c1 C, s]}_{bck}
% <c1, v, LeftP C c2,s >_{bck} &|-->& <c1(+)c2,left v, C>_{bck}
% {[c1 (+) c2, v', C, s ]}_{bck} &|-->& {[ c1, v, LeftP C c2, s[v' <><> left v] ]}_{bck}
% <c2, v, RightP c1 C, s>_{bck} &|-->& <c1(+)c2, right v, C>_{bck}
% {[c1 (+) c2, v', C,s ]}_{bck} &|-->& {[ c2, v, RightP c1 C, s[v' <><> right v] ]}_{bck}
% <c1, v1, LeftT C c2 v2, s>_{bck} &|-->& <c1(*)c2, (v1, v2), C, s>_{bck}
% <c2, v2, RightT c1 v1 C, s>_{bck} &|-->& {[ c1, v1, LeftT C c2 v2, s ]}_{bck}
% {[ c1 (*) c2, v, C, s ]}_{bck} &|-->& {[ c2, v2, RightT c1 v1 C, S' ]}_{bck}
% && where s' = s[v <><> (v1, v2)]

%subcode{opsem} include main
%! columnStyle = rclr
%! fwd = \triangleright
%! bck = \triangleleft
% <eps+, v, C, s>_{fwd} |--> <eps+, left (-v'), C, s[v <><> right v']>_{bck}
% <eps+, v, C, s>_{fwd} |--> <eps+, right v', C, s[v <><> left (-v')]>_{bck}

%subcode{opsem} include main
%! columnStyle = rclr
%! fwd = \triangleright
%! bck = \triangleleft
% {[eta+, v, C, s]}_{bck} |--> {[eta+, left (-v'), C, s[v <><> right v'] ]}_{fwd}
% {[eta+, v, C, s]}_{bck} |--> {[eta+, right v', C, s[v <><> left (-v')] ]}_{fwd}

>>>>

\dgm{if-c1-c2}

@@

As we will see if Sec. \ref{sec:neg-constructions}, every compact
closed category admits a {{trace}} operator. In the case of {{PiEE*}},
the operator {{trace*}} is defined only on the non-zero fragment.
Given {{f : a*c <-> b*c }}, we have {{trace* f : a <-> b}}:

A large class of constraint satisfaction problems can be expressed
using {{trace*}}. We illustrate the main ideas with the implementation
of a SAT-solver.

%% We proceed in small steps, reviewing some of the
%% necessary constructions presented in our earlier
%% paper~\cite{infeffects}.

\begin{comment}
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Finite Dimensional Hilbert Spaces}

\begin{definition}[Size of a type]
The defintion of size is extended to include negative types.
{{[^-b^] = -[^b^]}}
\end{definition}

%subcode{opsem} include main
%! columnStyle = rclr
% {[iso, v, C]}^^^ &|-->& <iso, v', C>^^^
% & & where iso^{dagger} v |--> v'
% <c1, v, Fst C c2>^^^ &|-->& <c1(;)c2, v, C>^^^
% <c2, v, Snd c1 C>^^^ &|-->& {[c1, v, Fst C c2]}^^^
% {[ c1(;)c2, v, C]}^^^ &|-->& {[c2, v, Snd c1 C]}^^^
% <c1, v, LeftP C c2>^^^ &|-->& <c1(+)c2,left v, C>^^^
% {[c1 (+) c2, left v, C]}^^^ &|-->& {[ c1, v, LeftP C c2 ]}^^^
% <c2, v, RightP c1 C, s>^^^ &|-->& <c1(+)c2, right v, C>^^^
% {[c1 (+) c2, right v, C ]}^^^ &|-->& {[ c2, v, RightP c1 C ]}^^^
% <c1, v1, LeftT C c2 v2>^^^ &|-->& <c1(*)c2, (v1, v2), C>^^^
% <c2, v2, RightT c1 v1 C >^^^ &|-->& {[ c1, v1, LeftT C c2 v2 ]}^^^
% {[ c1 (*) c2, (v1, v2), C ]}^^^ &|-->& {[ c2, v2, RightT c1 v1 C ]}^^^

%subcode{opsem} include main
%! columnStyle = rclr
%! fwd = \triangleright
%! bck = \triangleleft
% <eps+, right v, C, s> &|-->& <eps+, left (-v), C]>^^^
% <eps+, left (-v), C, s> &|-->& <eps+, right v, C>^^^

%subcode{opsem} include main
%! columnStyle = rclr
%! fwd = \triangleright
%! bck = \triangleleft
% {[eta+, right v, C, s]}^^^ &|-->& {[eta+, left (-v), C ]}
% {[eta+, left (-v), C, s]}^^^ &|-->& {[eta+, right v, C ]}

%%%%%%%%%
\subsection{Categorical Constructions}
\label{sec:neg-constructions}

\
All the constructions below are standard: they are collected from Selinger's
survey paper on monoidal
categories~\cite{springerlink:10.1007/978-3-642-12821-94} and presented in
the context of our language.

@@
